---
title: "Project 3: stat302package Tutorial"
author: "Yufei Wang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project 3: stat302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\addtolength{\headheight}{-.025\textheight} 
\thispagestyle{fancyplain} 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Vignette Information

We thank Bryan for teaching this amazing course Stat302 wit his passion and enthusiasm.

# Introduction
We develop a well-documented, well-tested, and well-explained R package for Project 3 of Stat302.This package includes thses functions which can be used to make inference and prediction based on dataset. Specifically, it includes four functions we’ve written throughout this quarter:
* my_t_test
* my_lm
* my_knn_cv
* my_rf_cv
Thus we use this package to do t-test, linear modeling, k-Nearest-Neighbors, and Random Forest Cross-Validation. 

Note that in order to follow along with this tutorial, you will need to installthe package from github by running the following code.
Install Package from Github repository:
```{r, eval = FALSE}
devtools::install_github("yufei-wang/stat302")
```

To begin, we load some necessary packages and dataset using:

```{r, message = FALSE}
library(stat302package)
library(stats)
library(magrittr)
library(dplyr)
library(class)
library(randomForest)
data("my_gapminder")
```

If you are unfamiliar with my_gapminder, make sure to check my_gapminder:

```{r}
?my_gapminder
```

## Tutorial

### Tutorial 1
The first tutorial is for my_t.test
