---
title: "Project 3: stat302package Tutorial"
author: "Yufei Wang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project 3: stat302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\addtolength{\headheight}{-.025\textheight} 
\thispagestyle{fancyplain} 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Vignette Information

We thank Bryan for teaching this amazing course Stat302 wit his passion and enthusiasm.

# Introduction
We develop a well-documented, well-tested, and well-explained R package for Project 3 of Stat302.This package includes thses functions which can be used to make inference and prediction based on dataset. Specifically, it includes four functions we’ve written throughout this quarter:
* my_t_test
* my_lm
* my_knn_cv
* my_rf_cv
Thus we use this package to do t-test, linear modeling, k-Nearest-Neighbors, and Random Forest Cross-Validation. 

Note that in order to follow along with this tutorial, you will need to installthe package from github by running the following code.
Install Package from Github repository:
```{r, eval = FALSE}
devtools::install_github("yufei-wang/stat302package")
```

To begin, we load some necessary packages and dataset using:

```{r, message = FALSE}
library(stat302package)
library(stats)
library(magrittr)
library(dplyr)
library(class)
library(randomForest)
library(kableExtra)
library(ggplot2)
data("my_gapminder")
```

If you are unfamiliar with my_gapminder, make sure to check my_gapminder:

```{r}
?my_gapminder
```

## Tutorial

### Tutorial 1:  my_t.test
The first tutorial is for my_t.test. We use `lifeExp` data from `my_gapminder` to do several t-tests.

#### Example 1: alternative = "two.sided"
$$H_0: \mu = 60$$
$$H_a: \mu \neq 60$$
$$\alpha = 0.05$$
```{r}
data("my_gapminder")
my_t.test(my_gapminder$lifeExp, alternative = "two.sided", mu = 60)
```
From the result, we fail to reject the null hypothesis at the 0.05 alpha level since p-value is greater than 0.05.

#### Example II: alternative = "less"
$$H_0: \mu = 60$$
$$H_a: \mu < 60$$
$$\alpha = 0.05$$
```{r}
my_t.test(my_gapminder$lifeExp, alternative = "less", mu = 60)
```
From the result, we reject the null hypothesis at the 0.05 alpha level since p-value is smaller than 0.05.

#### Example III: alternative = "greater"
$$H_0: \mu = 60$$
$$H_a: \mu > 60$$
$$\alpha = 0.05$$
```{r}
my_t.test(my_gapminder$lifeExp, alternative = "greater", mu = 60)
```
From the result, we fail to reject the null hypothesis at the 0.05 alpha level since p-value is greater than 0.05.


### Tutorial 2
The second tutorial is for my_lm. 
First, we demonstrate a regression using lifeExp as our response variable and gdpPercap and continent as explanatory variables:
```{r warning  = F}
tutor_model <- my_lm(lifeExp ~ gdpPercap + continent, data = my_gapminder)
tutor_model
```
Interpreting Coefficients: We would expect an average increase of 0.00045 for LifeExp when continent is unchanged.
Next, we will do the hypothesis test associated with the gdpPercap coefficient $\beta_1$ using the significant level of 0.05.
$H_0: \beta_1 = 0$
$H_0: \beta_1 \neq 0$
$ \alpha = 0.05$

From the result of the table, we reject the null hypothesis at the 0.05 alpha level since p-value is 0, which is smaller than 0.05.
Thus we could say that changes in gdpPercap is related to changes in lifeExp.

Then, we want to use ggplot2 to plot the Actual vs. Fitted values.
```{r}
coef <- tutor_model[, 1]
x_hat <- model.matrix(lifeExp ~ gdpPercap + continent, data = my_gapminder)
y_hat <- x_hat %*% as.matrix(coef)
my.df <- data.frame("Fitted" = y_hat, 
                    "Actual" = my_gapminder$lifeExp,
                    "Continent" = my_gapminder$continent)

ggplot(data = my.df, aes(x = Fitted, y = Actual, color = Continent)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  labs(x = "Fitted Values", y = "Actual Values", title = "Actual vs. Fitted Values",
       color = "Continent") +
  theme_bw(base_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))
```
Based on the plot, the line looks like only fit the data points of Europe and Oceania. The linear relationships indicates that the model is good for Europe and Oceania. However, for other continents, the model doesn't work properly. Thus, we should try other regressions.


### Tutorial 3
The third tutorial is for my_knn_cv. We will predict output class continent using covariates gdpPercap and lifeExp and 5-fold cross validation (k_cv = 5).
To do k-fold cross-validation, we split data into 5 parts (folds) at first. Then we use all but 1 fold as our training data and fit the model. After that, 
we use the remaining folds for test data and make predictions. In this way, we switch which fold is our test data and repeat this process until all folds 
have been test data. Finally, we compute squared error.

```{r}
train <- my_gapminder %>% select(gdpPercap, lifeExp)
cl <- my_gapminder$continent
k_cv = 5
#create empty matrix to save result
result.mat <- matrix(NA, 10, 2)
for (i in 1:10) {
  output <- my_knn_cv(train = train, cl = cl, k_nn = i, k_cv = k_cv)
  
  # record training misclassfication rate
  result.mat[i, 2] <- (sum(output$class != cl) / length(cl)) %>% round(5)
  
  # record CV misclassfication rate
  result.mat[i, 1] <- output$cv_err
  
}
result <- data.frame("Number of neighbors" = c(1:10), 
                     "CV misclassification rate" = result.mat[, 1],
                     "training misclassification rate" = result.mat[, 2])
kable_styling(kable(result))

```

Based on the result, if we are based on training misclassification rates, we will choose k_nn = 1 since the training misclassfication rate of k_nn = 1 is smallest, which is close to 0. If we are based on CV misclassification rates, we will choose k_nn = 10 since the cv error of k_nn = 10 is smallest among all of the knn value.

In practice, as a rule of thumb, we will choose k = 5 and k = 10 sincce these values tend to result in an ideal balance in terms of the bias-variance tradeoff.

 
### Tutorial 4
The fourth tutorial is for my_rf_cv.
```{r}
#create matrix to store cv error
cv_err.mat <- matrix(NA, 90, 2)
cv_err.mat[, 1] <- rep(c(2, 5, 10), 30)
j = 1
#For each value of k, run your function 30 times.
for(i in 1:30) {
  
    # Iterate through k in c(2, 5, 10):
    for(k in c(2, 5, 10)) {
      
    #For each of the 30 iterations, store the CV estimated MSE.
    cv_err.mat[j, 2] <- my_rf_cv(k) %>%  round(5)
    j = j +1
  }
}
cv.err <- data.frame("k" = cv_err.mat[, 1], "mse" = cv_err.mat[, 2])
ggplot(data = cv.err, aes(x = factor(k), y = mse), fill = factor(k)) +
  geom_boxplot(fill = "cornflowerblue") +
  labs(title = "Boxplots of MSE by k folds", x = "Number of Folds", y = "MSE") +
  theme_bw(base_size = 20) +
  theme(plot.title = element_text(hjust = 0.5))

table.df <- cv.err %>% group_by(k) %>% summarise("mean" = mean(mse), "sd" = sd(mse))
table.df %>% 
  kable("html") %>% 
  kable_styling()
```
From result of table, we observe that the standard error of MSE decreases as k increases. Meanwhile, the mean of MSE increases as k increases.
From result of boxplot, we can see that the range of MSE becomes narrow as k increases. Meanwhile, the median increases  as k increases.
It is because of bias-variance tradeoff. If we ncrease the complexity of a kind of model, the bias decreases and the variance increases. Thus, as k increases, the standard error of MSE decreases and mean of MSE increases.


